---
title: "STAT5003 Final Report (TUT10 Group6)"
author: |
  Sirui Li (530548597) · Marcus Yao (530688077) · Xi Huang(550902508) <br>
  Huiwen Jiang（510000080） · Kezhao Wu (540585108)
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: readable
    code_folding: hide 
---
<style>
body,
h1,h2,h3,h4,h5,h6,
h1.title, h4.author, h4.date {
  font-family: "Times New Roman", Times,
               "Noto Serif SC", "Source Han Serif SC", "Songti SC", "SimSun",
               serif;
}

body {
  font-size: 10px;   
  line-height: 1.5;
}

pre, code, kbd, samp {
  font-family: Consolas, "Courier New", ui-monospace, Menlo, Monaco, monospace;
  font-size: 0.95em;
}
</style>

<style>
h1 {
  font-size: 24px;
}
h2 {
  font-size: 20px;
}
h3 {
  font-size: 18px;
}
h1.title, h4.author, h4.date { text-align: center; }
.r-output, .r-message, .r-warning {
  max-height: 200px; overflow: auto; margin: .4rem 0 1rem; padding: .6rem;
  background: #f8f9fa; border-radius: 6px; border: 1px solid #e5e7eb;
}
pre.r { background: #f8f9fa; padding: .6rem; border-radius: 6px; }
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE
)
knitr::knit_hooks$set(output = function(x, options) {
  sprintf('<div class="r-output"><pre>%s</pre></div>', x)
})

# knitr::knit_hooks$set(message = function(x, options) sprintf('<div class="r-message"><pre>%s</pre></div>', x))
# knitr::knit_hooks$set(warning = function(x, options) sprintf('<div class="r-warning"><pre>%s</pre></div>', x))


```

# Problem Definition

For this task, we selected the Sydney dataset from the Inside Airbnb website. Our research question is: to predict the price category of a property based on its various features. This is a clear classification task, where the variable `price_category` has discrete categories—such as Low / Medium / High —as prices fall into distinct intervals. Our objective is to assign listings to appropriate price categories based on their attributes, making this a clear classification problem. This research holds significant societal value: as a major global tourist destination, Sydney benefits from our predictive model. Travelers can swiftly find budget-friendly accommodations, while hosts can optimize their listings by analyzing high-priced property characteristics. Overall, analyzing Sydney's Airbnb data provides valuable insights for travelers, governments, and tourism management agencies.

---

# Data Description
The dataset is Airbnb property information in Sydney which is obtained from [Inside Airbnb](http://insideairbnb.com/), which provides detailed information about Airbnb listings across Sydney, Australia. The dataset includes listing-level features such as location, property attributes, host information, and price. 
  
The dataset contains 18187 rows and 79 columns. Variables can be grouped into three categories:  
  - **Numerical variables:** price, minimum_nights, number_of_reviews, reviews_per_month, etc.  
  - **Categorical variables:** room_type, neighbourhood, host_is_superhost, etc.  
  - **Text variables:** listing name, description, host name, etc
  
**Outcome variable:**

The study segments listing **prices** into discrete intervals to serve as the categorical response. Models are then estimated using price-salient predictors to accurately classify and predict the price tiers of Sydney listings.

**Data Challenges**

During initial exploration, the following challenges are identified:
Missing values: Some variables (e.g., reviews_per_month) contain many missing entries.
Outliers: Price contains extreme values that may distort the distribution.
Class imbalance: Some categories (e.g., specific neighbourhoods or room types) dominate the dataset.
Text variables: Variables like description and name require NLP techniques if used

| Variable Type   | Examples                                  | Notes                                           |
| -------------------- | ---------------------------------------------- | -------------------------------------------------- |
| Numerical         | price, minimum\_nights, number\_of\_reviews    | Require cleaning and outlier handling /  |
| Categorical       | room\_type, neighbourhood, host\_is\_superhost | Require encoding /                             |
| Text             | name, description, host\_name                  | Optional features, may need NLP /     |
| Outcome          | price\_category (Low/Medium/High)              | Derived from price /                        |

These observations highlight both the richness and the challenges of the dataset. This naturally leads to the next step: Data Cleaning & Feature Engineering, where missing values, outliers, and categorical variables will be addressed to prepare the dataset for modeling.

---

# Data Cleaning & Feature Engineering
## Price Column Processing & Outlier Detection
We first preprocess the target column price by removing all non-numeric characters, converting price to numeric type, and removing observations with price ≤ 0. We also use IQR for outlier analysis.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)

# Read in the original data (all in character format to avoid misinterpretation of "price")
df <- read_csv("listings.csv", col_types = cols(.default = "c"))
```
**Step 1: Price Column Auto-Detection**
Target price by keyword, remove currency/thousands separators, and force conversion to number.
```{r, message=FALSE, warning=FALSE, results='hide'}
# Step 1: Find the price column
price_col <- NULL
possible_cols <- names(df)

# Check for 'price' (case-insensitive)
price_col <- possible_cols[str_detect(tolower(possible_cols), "price")][1]
if (is.na(price_col)) {
  # If not found, check for other keywords
  price_keywords <- c('cost', 'amount', 'value', 'fee', 'charge', 'payment')
  for (keyword in price_keywords) {
    found_col <- possible_cols[str_detect(tolower(possible_cols), keyword)][1]
    if (!is.na(found_col)) {
      price_col <- found_col
      break
    }
  }
}

if (is.na(price_col)) {
  stop("Error: Could not find a column representing price.")
}


# Cleaning comparison: Save the first 10 rows as the original, then generate the cleaned 10 rows, and combine and compare the tables.
price_raw10 <- head(df[[price_col]], 10)

df_clean_preview <- df %>%
  mutate(
    !!sym(price_col) := suppressWarnings(as.numeric(
      str_remove_all(
        str_remove_all(tolower(.data[[price_col]]), ","),
        "[^\\d.-]"
      )
    ))
  )

tibble(
  !!paste0(price_col, "_raw")   := price_raw10,
  !!paste0(price_col, "_clean") := head(df_clean_preview[[price_col]], 10)
) %>% knitr::kable(caption = "Price Comparison Before and After Cleaning (First 10 Lines)")

```
**Step 2: Price Standardization**
We normalize the price column by lowercasing it, removing commas and non-numeric characters, coercing it to numbers, and filtering out missing entries and prices ≤ 0.
```{r, message=FALSE, warning=FALSE}
# Step 2: Clean and convert the price column (corrected)

df <- df %>%
  mutate(
    # Remove commas and non-numeric characters, then convert to numeric.
    # `as.numeric` will correctly convert invalid values to `NA`.
    !!sym(price_col) := as.numeric(str_remove_all(str_remove_all(tolower(!!sym(price_col)), ","), "[^\\d.-]"))
  ) %>%
  # Now, filter out rows where the price is NA or <= 0
   dplyr::filter(!is.na(!!sym(price_col)) & !!sym(price_col) > 0)



```
**Step 3: IQR-Based Outlier Analysis for Price**
We conduct outlier detection on the cleaned price data using the Tukey IQR rule: we compute Q1, Q3, and IQR = Q3 - Q1, and define the inlier range as [Q1-1.5\*IQR, Q3+1.5\*IQR] (Lin, Zhang, & Tong, 2025). We then summarize key diagnostics in the table IQR-based Outlier Analysis Summary.(Dash, Behera, Dehuri, & Ghosh, 2023; Mekelburg & Strauss, 2024).
```{r, message=FALSE, warning=FALSE}
# Step 3: Analyze outliers using IQR (will now work correctly)
q1 <- quantile(df[[price_col]], 0.25)
q3 <- quantile(df[[price_col]], 0.75)
iqr <- q3 - q1
lower_bound <- q1 - 1.5 * iqr
upper_bound <- q3 + 1.5 * iqr

outliers <- df %>% dplyr::filter(!!sym(price_col) < lower_bound | !!sym(price_col) > upper_bound)


summary_tbl <- tibble::tibble(
  Metric = c(
    "Q1 (25th percentile)",
    "Q3 (75th percentile)",
    "Interquartile Range (IQR)",
    "Outlier Lower Bound",
    "Outlier Upper Bound",
    "Total Rows (N)",
    "Number of Outliers",
    "Percent of Data"
  ),
  Value = c(
    round(q1, 2),
    round(q3, 2),
    round(iqr, 2),
    round(lower_bound, 2),
    round(upper_bound, 2),
    nrow(df),
    nrow(outliers),
    paste0(round(nrow(outliers) / max(1, nrow(df)) * 100, 2), "%")
  )
)

knitr::kable(summary_tbl, caption = "IQR-based Outlier Analysis Summary")
```
We detect outliers with Tukey’s IQR rule and summarize Q1, median, IQR, and fences. Instead of dropping rows, we winsorize the price column: values below P3 (55) are set to 55 and above P97 (1350) to 1350, keeping N = 15,735. About 920 values (≈5.9%) were clipped across tails. This reduces extreme influence while preserving distributional shape for robust analysis (Dash et al., 2023; Mekelburg & Strauss, 2024; Lin et al., 2025).
```{r, message=FALSE, warning=FALSE, results='hide'}
stopifnot(exists("df"), exists("price_col"))

# Calculate 3rd and 97th percentiles for winsorization
p_lo <- quantile(df[[price_col]], 0.03, na.rm = TRUE)
p_hi <- quantile(df[[price_col]], 0.97, na.rm = TRUE)

# Winsorize the price column and track clipping   
df <- df %>%
  dplyr::mutate(
    !!rlang::sym(paste0(price_col, "_orig")) := .data[[price_col]],
    !!rlang::sym(price_col) := pmin(pmax(.data[[price_col]], p_lo), p_hi),
    .clipped_low  = .data[[paste0(price_col, "_orig")]] < p_lo,
    .clipped_high = .data[[paste0(price_col, "_orig")]] > p_hi
  )

# Summarize clipping
summary_tbl <- tibble::tibble(
  P1  = round(p_lo,  2),
  P99 = round(p_hi,  2),
  N   = nrow(df),
  Clipped_Low   = sum(df$.clipped_low,  na.rm = TRUE),
  Clipped_High  = sum(df$.clipped_high, na.rm = TRUE),
  Total_Clipped = sum(df$.clipped_low | df$.clipped_high, na.rm = TRUE),
  Percent_Clipped = paste0(
    round(100 * mean(df$.clipped_low | df$.clipped_high, na.rm = TRUE), 2), "%"
  )
)
knitr::kable(summary_tbl, caption = "Winsorization Summary (3%–97% percentiles)")


```
**Step 4: Save cleaned data**
After completing the above cleaning steps, if necessary, the cleaned data can be saved into a new file named "cleaned_data.csv" by following code.

```{r}
# Step 4: Save the further cleaned data
write_csv(df, "cleaned_data.csv")
```
## Feature Engineering: Price Categorization
Building on the cleaned data, we re-detect the price column, coerce it to numeric, and bin prices into three tiers—[0,150), [150,300), and [300,+)—with non-positive or missing values labeled Invalid Price.
```{r, message=FALSE, warning=FALSE, results='hide'}
# Re-read the cleaned data from the previous step
df <- read_csv("cleaned_data.csv", col_types = cols(.default = "c"))

# Find the price column again
price_col <- names(df)[str_detect(tolower(names(df)), "price")][1]

# Check if the price column was found
if (is.na(price_col)) {
  stop("Error: Could not find a column representing price.")
}

# Step 1: Ensure the price column is numeric before creating 'price_class'
# This is a crucial step to prevent errors
df <- df %>%
  mutate(
    # Safely convert the price column to numeric,
    # invalid parsing will result in NA
    price_numeric = as.numeric(!!sym(price_col)),
    
    # Create the new 'price_class' column based on the numeric price
    price_class = case_when(
      is.na(price_numeric) | price_numeric <= 0 ~ "Invalid Price",
      price_numeric < 150 ~ "[0,150)",
      price_numeric < 300 ~ "[150,300)",
      TRUE ~ "[300,+)"
    )
  )

# Step 2: Reorder columns to place 'price_class' at the beginning
df <- dplyr::select(df, price_class, dplyr::everything(), -price_numeric)

# Step 3: Save the data with the new column
write_csv(df, "cleaned_data_with_price_class.csv")

```

## Missing Value Imputation
We detected missing values and classified columns as numeric or non-numeric via an 80% coercion rule. Numeric NAs were replaced with -1, non-numeric with "NA", and results saved as cleaned_data_with_price_class_no_na.csv. We tracked missingness by computing per-column flag rates (share of -1 for numeric, "NA"/"-1" for non-numeric), preserving information for downstream modeling.
```{r, message=FALSE, warning=FALSE}
# Load necessary libraries
# You might need to install these packages first: install.packages("tidyverse")
library(readr)
library(dplyr)
library(ggplot2)
library(stringr)

# Re-read the data
# suppressWarnings is added here as `read_csv` might generate warnings
df <- suppressWarnings(read_csv("cleaned_data_with_price_class.csv", col_types = cols(.default = "c")))

# Step 1: Identify numeric and non-numeric columns using the Python logic
numeric_cols <- character(0)
non_numeric_cols <- character(0)

for (col in names(df)) {
  sample <- df[[col]][!is.na(df[[col]]) & df[[col]] != "NA"][1:100]
  if (length(sample) > 0) {
    # suppressWarnings is used here to hide the 'NAs introduced by coercion' warnings
    numeric_count <- suppressWarnings(sum(!is.na(as.numeric(sample))))
    if (numeric_count / length(sample) > 0.8) {
      numeric_cols <- c(numeric_cols, col)
    } else {
      non_numeric_cols <- c(non_numeric_cols, col)
    }
  } else {
    non_numeric_cols <- c(non_numeric_cols, col)
  }
}

# Step 2: Impute missing values
df_imputed <- df %>%
  mutate(
    # Impute NA in numeric columns with -1
    across(all_of(numeric_cols), ~replace_na(as.numeric(.x), -1)),
    # Impute NA in non-numeric columns with 'NA'
    across(all_of(non_numeric_cols), ~replace_na(.x, 'NA'))
  )

# Step 3: Save the final cleaned data
write_csv(df_imputed, "cleaned_data_with_price_class_no_na.csv")

# Step 4: Create a function to calculate missing rates
calculate_missing_rates <- function(data) {
  missing_rates <- sapply(names(data), function(col) {
    # Check for -1 in columns that have been correctly converted to numeric
    if (is.numeric(data[[col]])) {
      sum(data[[col]] == -1) / nrow(data)
    } else {
      # Check for 'NA' or '-1' in non-numeric columns
      sum(data[[col]] %in% c('NA', '-1')) / nrow(data)
    }
  })
  sort(missing_rates * 100, decreasing = TRUE)
}


missing_rates <- calculate_missing_rates(df_imputed)
print(missing_rates)
```
The following code can display the ten columns with the most missing values in the data, and the statistical significant result of its improvement.
```{r, message=FALSE, warning=FALSE, results='hide'}

library(readr); library(dplyr); library(knitr)

df_before <- read_csv("cleaned_data_with_price_class.csv",
                      col_types = cols(.default = "c"), show_col_types = FALSE)
df_after  <- read_csv("cleaned_data_with_price_class_no_na.csv",
                      show_col_types = FALSE)

n <- nrow(df_before)

count_before <- function(x) if (is.character(x)) sum(is.na(x) | x == "" | x == "NA") else sum(is.na(x))
count_after <- function(x) {
  if (is.numeric(x)) {
    sum(x == -1, na.rm = TRUE)              
  } else if (is.character(x)) {
    sum(x %in% c("NA", ""), na.rm = TRUE)   
  } else {
    sum(is.na(x))                            
  }
}

before_n <- sapply(df_before, count_before)
after_n  <- sapply(df_after,  count_after)

comp_tbl <- tibble::tibble(
  column        = names(df_before),
  before_n      = as.integer(before_n),
  before_pct    = round(100 * before_n / n, 2),
  after_n       = as.integer(after_n),
  after_pct     = round(100 * after_n / n, 2),
  reduction_n   = before_n - after_n,
  reduction_pct = round(100 * (before_n - after_n) / n, 2)
) %>% arrange(desc(reduction_n))

# Reduce the number of printed columns by the greatest extent.
print(kable(head(comp_tbl, 10), caption = "The 10 columns with the greatest reduction in missing/marked data"))

# Overall Comparison (Full Table)
overall <- tibble::tibble(
  before_n      = sum(before_n),
  before_pct    = round(100 * sum(before_n) / (n * ncol(df_before)), 2),
  after_n       = sum(after_n),
  after_pct     = round(100 * sum(after_n) / (n * ncol(df_after)),  2),
  reduction_n   = sum(before_n) - sum(after_n),
  reduction_pct = round(100 * (sum(before_n) - sum(after_n)) / (n * ncol(df_before)), 2)
)
print(kable(overall, caption = "Overall absence/label comparison"))
```

## Post-processing and Feature Curation
After completing the data cleaning, we still have some subsequent processing to do on the data. The following steps streamline the dataset, reduce noise introduced by poorly populated or non-analytical fields, and standardize key attributes for reliable EDA and modeling.
**Column order adjustment: **
Reordered columns so that id precedes price_class to improve identifier visibility during analysis.
**Location normalization: **
Atomicized host_location into the most granular, indivisible values to reduce ambiguity and enable consistent grouping/joining.
**Row integrity filtering: **
Removed records with column misalignment detected during manual inspection to prevent downstream parsing errors.
**Low-value / high-missingness features removed:**
    * `listing_url` * `scrape_id` * `description` * `neighborhood_overview` * `picture_url`
    * `host_id` * `host_url` * `host_name` * `host_about` * `host_thumbnail_url` * `host_picture_url`
    * `neighbourhood_group_cleansed` * `calendar_updated`
These features were removed either because they had little analytical value or because their missing data rate was close to 100%.
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(stringr)
library(readr)
library(lubridate)

df_filtered <- df_after %>%
  dplyr::select(-c(
    listing_url, scrape_id, description, neighborhood_overview, picture_url,
    host_id, host_url, host_name, host_about, host_thumbnail_url, host_picture_url,
    neighbourhood_group_cleansed, calendar_updated
  ))


df_filtered <- df_filtered %>%
  mutate(
    host_response_rate = as.character(host_response_rate),
    host_response_rate = str_trim(host_response_rate),
    host_response_rate = ifelse(host_response_rate == "" | host_response_rate == "N/A", NA, host_response_rate)
  )


df_filtered <- df_filtered %>%
  dplyr::filter(
    is.na(host_response_rate) | 
    str_detect(host_response_rate, "^(100|[0-9]{1,2})%$")
  )

df_filtered <- df_filtered %>%
  mutate(
    host_since_char = as.character(host_since),
    host_since_char = str_trim(host_since_char)
  )

problematic_entries <- df_filtered %>%
  mutate(
    has_letters = str_detect(host_since_char, "[a-zA-Z]"),
    has_quotes = str_detect(host_since_char, '[""]'),
    is_numeric_date_like = str_detect(host_since_char, "^[0-9/-]+$"),
    char_length = nchar(host_since_char)
  ) %>%
  dplyr::filter(!is.na(host_since_char))

entries_with_letters <- problematic_entries %>%
  dplyr::filter(has_letters) %>%
  dplyr::select(host_since_char, char_length) %>%
  distinct() %>%
  arrange(char_length)

long_entries <- problematic_entries %>%
  dplyr::filter(char_length > 20) %>%
  dplyr::select(host_since_char, char_length) %>%
  distinct() %>%
  arrange(desc(char_length))

numeric_entries <- problematic_entries %>%
  dplyr::filter(is_numeric_date_like & char_length <= 15) %>%
  dplyr::select(host_since_char) %>%
  distinct() %>%
  head(20)

write_csv(df_filtered, "cleaned_data_with_price_class_no_dislocation.csv")

# Display final data overview
print(dim(df_filtered))
```
After executing the above code, we first removed irrelevant columns to simplify the dataset. Then, we cleaned the host_response_rate column by trimming spaces, converting empty strings and "N/A" values into missing values, and keeping only valid percentages between 0% and 100%. Next, we processed the host_since column by converting it to character format and checking for problematic entries.

# Exploratory Data Analysis (EDA)

## Distribution of the Target Variable
The target variable price_class is divided into three ranges: low [0,150), medium [150,300), and high [300,+). As shown in the bar chart, the medium price category contains the largest number of listings, with more than 6,500 entries. The low and high price categories are close in size, each with around 4,500 listings. This visualization provides an overview of how listings are distributed across price ranges, setting the stage for further analysis in subsequent sections.
```{r,out.width="50%",fig.align="center",out.height="50%"}
library(patchwork)
library(corrplot)
library(ggplot2)
library(tidyverse)
df <- read.csv("cleaned_data_with_price_class_no_dislocation.csv", encoding = "latin1")

# Use a bar chart to visualize the distribution of price_class.
ggplot(df, aes(x = price_class)) +
  geom_bar(fill = "steelblue") +
  theme_minimal() +
  labs(title = "Distribution of Target Variable: price_class")
```

## Categorical feature

### Room Type vs. Price Class
The distribution of room_type across price_class categories shows clear differences. Entire home/apt dominates the dataset, especially in the medium and high price ranges. Private rooms are mostly concentrated in the low-price category, while hotel rooms and shared rooms are rare and contribute minimally. This pattern highlights room type as a strong determinant of listing price, with full-property rentals generally commanding higher values.
```{r,out.width="50%",fig.align="center"}
# Room Type vs Price Class
  ggplot(df, aes(x = room_type, fill = price_class)) +
    geom_bar(position = "dodge") +
    theme_minimal() +
    labs(title = "Room Type vs Price Class")
```

### Neighbourhood vs. Price Class 
The distribution of listings across the top 10 neighbourhoods shows notable variation in price categories. Sydney dominates the dataset, with a majority of listings in the medium and high price classes. In contrast, areas such as Randwick, and North Sydney present a more balanced mix of low- and mid-priced properties, while suburbs like Waverley and Manly display relatively higher proportions of high-price listings. These differences highlight the strong influence of location on pricing, confirming that neighbourhood is a key determinant of property value alongside room type.

```{r,out.width="50%", fig.align="center"}
# Step 1: Calculate the frequency and select the top ten
top_10 <- names(sort(table(df$neighbourhood_cleansed), decreasing = TRUE))[1:10]

# Step 2: Filtering data
df_subset <- df[df$neighbourhood_cleansed %in% top_10, ]

# Step 3: Visualize
ggplot(df_subset, aes(x = neighbourhood_cleansed, fill = price_class)) +
  geom_bar(position = "dodge") +
  labs(title = "Top 10 Neighbourhoods vs Price Class",
       x = "Neighbourhood",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## Numerical feature

The feature accommodates was selected for exploratory analysis as it directly reflects a property’s capacity and is expected to influence price. Correlation analysis further showed that accommodates, bedrooms, and beds are strongly related and positively associated with price_class. Therefore, accommodates was selected as a representative feature to examine how property’s capacity contributes to price segmentation.

### Accommodates vs. Price Class
The distribution of accommodates is highly right-skewed, with most listings designed to host between 1 and 4 guests, indicating that smaller properties dominate the dataset.
The boxplot comparison across price_class shows a clear increasing pattern: higher price categories generally correspond to listings that can accommodate more guests.
The median capacity rises from around 2 in the lowest price class to about 6 in the highest, reflecting the intuitive link between property size and price level.
This visualization highlights that capacity is an important numerical feature to consider in modeling.

```{r, fig.width=10, fig.height=5, out.width="70%", fig.align="center"}
# 1. Histogram: Distribution of Accommodates
p1 <- ggplot(df, aes(x = accommodates)) +
  geom_histogram(binwidth = 1,fill = "skyblue", color = "black") +
  labs(title = "Distribution of Accommodates",
       x = "Number of Accommodates",
       y = "Count")

# 2. Box plot: Accommodates vs Price Class
p2 <- ggplot(df, aes(x = price_class, y = accommodates, fill = price_class)) +
  geom_boxplot() +
  labs(title = "Accommodates vs Price Class",
       x = "Price Class",
       y = "Accommodates") +
  theme(legend.position = "none")

p1 + p2  
```

## Correlation Heatmap of Numerical Features 

To improve readability and focus on the most relevant insights, we selected the top 12 numerical features most strongly correlated with the target variable price_class. This approach highlights the variables that are most likely to influence price segmentation while avoiding excessive visual clutter from weakly related features. The resulting heatmap provides a clear view of key relationships and helps identify potential multicollinearity among the most influential predictors.
```{r,out.width="70%", fig.align="center"}

df <- df %>% 
  mutate(price_class_numeric = case_when(
    price_class == "[0,150)" ~ 1,
    price_class == "[150,300)" ~ 2,
    price_class == "[300,+)" ~ 3,
    TRUE ~ NA_real_
  ))

# Step 1: Selection of numerical features
numeric_features <- df %>%
  select(where(is.numeric)) %>%
  select(-price_class_numeric) %>% # Excluding the numerical version of the target variable
  select(-price) %>%
  select(-price_orig)

# Step 2: Calculate the correlation of each numerical feature with the target variable
target_correlations <- numeric_features %>%
  map_dbl(~ cor(., df$price_class_numeric, use = "complete.obs")) %>%
  abs() %>%  # Take absolute values and focus on strength of correlation rather than direction
  sort(decreasing = TRUE) %>%
  head(12)   # Select the 12 features with the highest relevance


selected_features <- names(target_correlations)

# Calculate the matrix of correlation coefficients between these features
cor_matrix <- cor(numeric_features[selected_features], use = "complete.obs")

# Visualisation of heat maps
par(mar = c(0, 0, 0, 0) + 0.1) 
corrplot(cor_matrix,
         method = "square",
         type = "upper",
         tl.cex = 0.6,
         tl.col = "darkblue",
         tl.srt = 45,
         addCoef.col = "black",
         number.cex = 0.4,
         cex.main = 1.2,
         col = colorRampPalette(c("blue", "white", "red"))(100),
        )

title("Correlation Heatmap of Top 12 Numerical Features", 
      line = -2, 
      adj = 0,   
      cex.main = 1.0)

```
The heatmap reveals strong positive correlations among accommodates, bedrooms, and beds, suggesting potential multicollinearity. Other numerical features show weaker correlations and may provide complementary information for modeling.

## Class Imbalance Check

We examined the distribution of the target variable (price_class) and two key categorical features (room_type, neighbourhood_cleansed).  
Price Class: Moderately imbalanced, with most listings in the 150–300 range. All classes remain reasonably represented.  
Room Type: Strongly skewed toward entire home/apartment; private rooms are fewer, hotel/shared rooms are rare.  
Neighbourhood: Highly concentrated, with Sydney dominating; other neighbourhoods have fewer listings.  
Overall, class imbalance is moderate for price_class but significant for room_type and neighbourhood_cleansed, which may require weighting or stratified sampling in modeling.

```{r}
# Visualization: bar charts
p1 <- ggplot(df, aes(x = price_class)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Price Class Distribution", x = "Price Class", y = "Count") +
  theme_minimal()

p2 <- ggplot(df, aes(x = room_type)) +
  geom_bar(fill = "tomato") +
  labs(title = "Room Type Distribution", x = "Room Type", y = "Count") +
  theme_minimal()

# Only the top 10 most frequent Neighborhoods are shown.
top_neigh <- names(sort(table(df$neighbourhood_cleansed), decreasing = TRUE))[1:10]
df_top <- df[df$neighbourhood_cleansed %in% top_neigh, ]

p3 <- ggplot(df_top, aes(x = neighbourhood_cleansed)) +
  geom_bar(fill = "darkseagreen") +
  labs(title = "Top 10 Neighbourhood Distribution", x = "Neighbourhood", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

(p1 | p2) / p3

```

---------





# Classification Model Used
## Softmax Logistic Regression
### Model Overview
Softmax logistic regression (multinomial logistic regression) is selected as the baseline model for this multi-class classification task. The model estimates the probability of each price class using the softmax function, providing interpretable coefficients that indicate the direction and magnitude of each feature's effect on class membership.


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(nnet)         # For multinom()
library(caret)        # For confusionMatrix and createFolds
library(pROC)         # For ROC curves
library(MLmetrics)    # For additional metrics
```

### Data Loading and Initial Preparation
The analysis begins by loading the cleaned dataset from previous preprocessing steps. Invalid price entries are removed, and the target variable `price_class` is converted to an ordered factor with three levels: low [0,150), medium [150,300), and high [300,+).

```{r, message=FALSE, warning=FALSE}
# Load cleaned data from previous preprocessing steps
df <- read.csv("cleaned_data_with_price_class_no_dislocation.csv", 
               encoding = "latin1")

cat("Original dimensions:", nrow(df), "rows ×", ncol(df), "columns\n")

# Remove invalid rows and normalize target labels (keep original style)
df <- df %>%
  dplyr::filter(price_class != "Invalid Price") %>%
  mutate(
    # remove any stray spaces: e.g. "[300, +)" -> "[300,+)"
    price_class = gsub("\\s+", "", price_class),
    # ensure only the three canonical labels are kept
    price_class = ifelse(price_class %in% c("[0,150)", "[150,300)", "[300,+)"),
                         price_class, NA_character_),
    price_class = factor(price_class,
                         levels = c("[0,150)", "[150,300)", "[300,+)"),
                         ordered = TRUE)
  ) %>%
  tidyr::drop_na(price_class)

# optional: check levels
print(levels(df$price_class))
```

### Feature Selection Strategy
Features are selected based on their analytical value and data quality. Non-informative columns (IDs, URLs, text fields) and date variables are excluded. Numeric features with more than 50% missing values are filtered out to ensure model stability. Key categorical features with reasonable cardinality are retained for modeling.

```{r, message=FALSE, warning=FALSE}
# Define columns to exclude from modeling
exclude_cols <- c(
  "id", "price_class",  # ID and target variable
  "price", "price_orig",  # Original price columns
  "name", "host_location",  # Unstructured text
  "latitude", "longitude",  # Spatial coordinates
  "last_scraped", "host_since", "first_review", "last_review",  # Date variables
  "neighbourhood"  # Redundant with neighbourhood_cleansed
)

# Select features for modeling
modeling_features <- setdiff(names(df), exclude_cols)

# Identify numeric and categorical features
numeric_cols <- df %>% 
  select(all_of(modeling_features)) %>% 
  select(where(is.numeric)) %>% 
  names()

categorical_cols <- df %>% 
  select(all_of(modeling_features)) %>% 
  select(where(~is.character(.) || is.factor(.))) %>% 
  names()

cat("Numeric features:", length(numeric_cols), "\n")
cat("Categorical features:", length(categorical_cols), "\n")

# Filter numeric features with high missingness (>50%)
valid_numeric <- numeric_cols[sapply(numeric_cols, function(col) {
  x <- df[[col]]
  missing_rate <- sum(x == -1 | is.na(x)) / length(x)
  missing_rate < 0.5
})]

cat("Valid numeric features after filtering:", length(valid_numeric), "\n")

# Select key categorical features
key_categorical <- c("room_type", "neighbourhood_cleansed", 
                     "host_is_superhost", "instant_bookable")
key_categorical <- key_categorical[key_categorical %in% categorical_cols]

cat("Key categorical features:", length(key_categorical), "\n")
```

### Preprocessing Pipeline
A preprocessing function is defined to handle data transformations within each cross-validation fold, preventing data leakage. The function performs two main tasks: (1) replacing missing value indicators (-1) with the median computed from the training set, and (2) one-hot encoding categorical variables while limiting high-cardinality features to their top 10 most frequent categories.

```{r, message=FALSE, warning=FALSE}
preprocess_for_modeling <- function(train_df, test_df, 
                                    num_features, cat_features) {
  
  # --- Process Numeric Features ---
  X_train_num <- train_df[, num_features, drop = FALSE]
  X_test_num <- test_df[, num_features, drop = FALSE]
  
  # Replace -1 (missing indicator) with median from training set
  for (col in num_features) {
    train_vals <- X_train_num[[col]]
    valid_vals <- train_vals[train_vals != -1 & !is.na(train_vals)]
    
    if (length(valid_vals) > 0) {
      col_median <- median(valid_vals)
      X_train_num[[col]][X_train_num[[col]] == -1] <- col_median
      X_test_num[[col]][X_test_num[[col]] == -1] <- col_median
    }
  }
  
  # Note: Standardization is not applied to preserve interpretability
  
  # --- Process Categorical Features (One-Hot Encoding) ---
  X_train_cat <- data.frame(row.names = 1:nrow(train_df))
  X_test_cat <- data.frame(row.names = 1:nrow(test_df))
  
  for (col in cat_features) {
    train_vals <- as.character(train_df[[col]])
    test_vals <- as.character(test_df[[col]])
    
    # Get unique values from training set (excluding "NA")
    unique_vals <- unique(train_vals)
    unique_vals <- unique_vals[unique_vals != "NA" & !is.na(unique_vals)]
    
    # Limit to top 10 categories for high-cardinality features
    if (length(unique_vals) > 10) {
      top_10 <- names(sort(table(train_vals), decreasing = TRUE)[1:10])
      train_vals <- ifelse(train_vals %in% top_10, train_vals, "Other")
      test_vals <- ifelse(test_vals %in% top_10, test_vals, "Other")
      unique_vals <- c(top_10, "Other")
    }
    
    # Create dummy variables (drop first level to avoid multicollinearity)
    if (length(unique_vals) > 1) {
      for (val in unique_vals[-1]) {
        dummy_col <- paste0(col, "_", make.names(val))
        X_train_cat[[dummy_col]] <- as.numeric(train_vals == val)
        X_test_cat[[dummy_col]] <- as.numeric(test_vals == val)
      }
    }
  }
  
  # Combine numeric and categorical features
  X_train_final <- cbind(X_train_num, X_train_cat)
  X_test_final <- cbind(X_test_num, X_test_cat)
  
  # Ensure test set has same columns as training set
  missing_cols <- setdiff(names(X_train_final), names(X_test_final))
  for (col in missing_cols) {
    X_test_final[[col]] <- 0
  }
  X_test_final <- X_test_final[, names(X_train_final)]
  
  return(list(
    X_train = X_train_final,
    y_train = train_df$price_class,
    X_test = X_test_final,
    y_test = test_df$price_class
  ))
}
```

### Model Training with Cross-Validation
To obtain robust performance estimates, 5-fold stratified cross-validation is employed. In each fold, the dataset is split into training and test sets while maintaining class proportions. Preprocessing is applied independently within each fold to prevent information leakage. The multinomial logistic regression model is fitted using the `nnet::multinom()` function with a maximum of 200 iterations.

```{r, message=FALSE, warning=FALSE}
set.seed(123)
k_folds <- 5
folds <- createFolds(df$price_class, k = k_folds, list = TRUE)

# Initialize storage for cross-validation results
cv_results <- data.frame(
  fold = integer(),
  accuracy = numeric(),
  macro_f1 = numeric(),
  mcc = numeric(),
  macro_auc = numeric(),
  logloss = numeric()
)

# Store predictions for later analysis
all_predictions <- list()
all_true_labels <- list()

cat("\nPerforming 5-fold stratified cross-validation...\n")

for (fold_i in 1:k_folds) {
  cat(sprintf("\nFold %d/%d: ", fold_i, k_folds))
  
  # Create train-test split for current fold
  test_idx <- folds[[fold_i]]
  train_fold <- df[-test_idx, ]
  test_fold <- df[test_idx, ]
  
  # Apply preprocessing within fold
  prep <- preprocess_for_modeling(train_fold, test_fold, 
                                   valid_numeric, key_categorical)
  
  # Train model on fold's training data
train_fold_data <- cbind(target = prep$y_train, prep$X_train)

model_cv <- tryCatch({
  multinom(target ~ ., data = train_fold_data, 
           maxit = 200, trace = FALSE, MaxNWts = 10000)
}, error = function(e) {
  cat("ERROR -", e$message, "\n")
  return(NULL)
})

if (is.null(model_cv)) next

# Generate predictions on fold's test data
pred_prob_cv <- predict(model_cv, newdata = prep$X_test, type = "probs")
pred_class_cv <- predict(model_cv, newdata = prep$X_test, type = "class")

# Store predictions for aggregation
all_predictions[[fold_i]] <- pred_class_cv
all_true_labels[[fold_i]] <- prep$y_test

# Evaluate fold performance
cm_cv <- confusionMatrix(pred_class_cv, prep$y_test)

acc_cv <- cm_cv$overall['Accuracy']
f1_cv <- mean(cm_cv$byClass[, "F1"], na.rm = TRUE)

# Calculate MCC for fold
n_cv <- sum(cm_cv$table)
correct_cv <- sum(diag(cm_cv$table))
p_cv <- rowSums(cm_cv$table)
q_cv <- colSums(cm_cv$table)
num_cv <- n_cv * correct_cv - sum(p_cv * q_cv)
denom_cv <- sqrt((n_cv^2 - sum(q_cv^2)) * (n_cv^2 - sum(p_cv^2)))
mcc_cv <- ifelse(denom_cv == 0, 0, num_cv / denom_cv)

# Calculate AUC for fold
auc_cv <- numeric(3)
for (i in 1:3) {
  y_bin <- as.numeric(prep$y_test == levels(prep$y_test)[i])
  if (is.matrix(pred_prob_cv)) {
    prob_i <- pred_prob_cv[, i]
  } else {
    prob_i <- ifelse(pred_class_cv == levels(prep$y_test)[i], 1, 0)
  }
  roc_cv <- roc(y_bin, prob_i, quiet = TRUE)
  auc_cv[i] <- auc(roc_cv)
}
macro_auc_cv <- mean(auc_cv)

# Calculate Log Loss for fold
ll_cv <- MultiLogLoss(y_pred = pred_prob_cv, y_true = prep$y_test)

# Store fold results
cv_results <- rbind(cv_results, data.frame(
  fold = fold_i,
  accuracy = acc_cv,
  macro_f1 = f1_cv,
  mcc = mcc_cv,
  macro_auc = macro_auc_cv,
  logloss = ll_cv
))

cat(sprintf("Acc=%.4f, F1=%.4f, MCC=%.4f\n", acc_cv, f1_cv, mcc_cv))
}

cat("\n5-fold cross-validation completed.\n")
```



## Linear Discriminant Analysis (LDA)
Linear Discriminant Analysis is a supervised classification problem that helps separate two or more classes by converting higher-dimensional data space into a lower-dimensional space. It is used to identify a linear combination of features that best separates classes within a dataset.

Therefore, by testing and comparing LDA on this dataset, we can validate LDA’s data-assumption requirements and identify other classification methods that may achieve better performance on this dataset.

### Extract Key Feature and Drop NA Value
From the previous observation in 'Class Imbalance Check', we know that the two key categorical features are room_type and neighbourhood_cleansed, so we extract the key features first before starting our LDA process.
```{r, message=FALSE, warning=FALSE, results='hide'}
library(readr)
library(dplyr)
library(tidyr)

file <- "cleaned_data_with_price_class_no_dislocation.csv"
target_cols <- c("price_class", "room_type", "neighbourhood_cleansed")

df_selected <- read_csv(
  file,
  col_select = all_of(target_cols)
)

df_lda <- df_selected %>%
    mutate(
        price_class = factor(price_class),
        room_type = factor(room_type),
        neighbourhood_cleansed = factor(neighbourhood_cleansed)
    )


df_after <- df_lda %>%
    drop_na(price_class, room_type, neighbourhood_cleansed)

```
### LDA Process
```{r, message=FALSE, warning=FALSE, results='hide'}
library(MASS)
library(conflicted)
conflict_prefer("select", "dplyr")
library(ggplot2)

lda_fit <- lda(price_class ~ room_type + neighbourhood_cleansed, data = df_after)
lda_pred <- predict(lda_fit, newdata = df_after)

# Confusion Matrix
tab <- table(Predicted = lda_pred$class, True = df_after$price_class)
acc <- sum(diag(tab)) / sum(tab)

# Visualization
ld_df <- data.frame(lda_pred$x, price_class = df_after$price_class)
ggplot(ld_df, aes(x = LD1, y = LD2, color = price_class)) +
  geom_point(size = 2, alpha = 0.7) +
  theme_minimal() +
  labs(title = "LDA projection (LD1 vs LD2)",
       x = "LD1", y = "LD2")


```

```{r}
library(yardstick)
library(conflicted)


# helper functions for multi-class MCC
mcc_binary <- function(t, p) {
  tp <- sum(p == 1 & t == 1)
  tn <- sum(p == 0 & t == 0)
  fp <- sum(p == 1 & t == 0)
  fn <- sum(p == 0 & t == 1)
  denom <- sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
  if (is.na(denom) || denom == 0) return(0)
  (tp * tn - fp * fn) / denom
}

mcc_macro <- function(truth, pred) {
  lv <- levels(truth)
  mccs <- sapply(lv, function(l) {
    t <- ifelse(truth == l, 1, 0)
    p <- ifelse(pred == l, 1, 0)
    mcc_binary(t, p)
  })
  mean(mccs, na.rm = TRUE)
}

# Compute LDA performance metrics
lda_results <- data.frame(
  truth = df_after$price_class,
  estimate = lda_pred$class
)

lda_acc <- mean(lda_results$truth == lda_results$estimate)
lda_prec <- yardstick::precision(lda_results, truth = truth, estimate = estimate, estimator = "macro")$.estimate
lda_rec  <- yardstick::recall(lda_results, truth = truth, estimate = estimate, estimator = "macro")$.estimate
lda_f1   <- yardstick::f_meas(lda_results, truth = truth, estimate = estimate, estimator = "macro")$.estimate
lda_mcc  <- mcc_macro(lda_results$truth, lda_results$estimate)

# Print summary
cat("\nLDA MODEL PERFORMANCE SUMMARY:\n")
cat(sprintf("  Accuracy:  %.4f\n", lda_acc))
cat(sprintf("  Precision (Macro): %.4f\n", lda_prec))
cat(sprintf("  Recall (Macro):    %.4f\n", lda_rec))
cat(sprintf("  F1 Score (Macro):  %.4f\n", lda_f1))
cat(sprintf("  MCC:               %.4f\n", lda_mcc))
```

Then we got the result of LDA with a visualized plot, which shows the relation between LD1 and LD2, regarding to the 'price_class'. The detailed evaluation of LDA will be discussed later in the chapter of 'Performance Evaluation'. Next, let's dive into the implementation of KNN.

## K Nearest Neighbour (kNN)
The k-Nearest Neighbour (kNN) algorithm is a non-parametric, instance-based learning method that classifies a new sample by the majority class of its k nearest neighbours in the feature space. It does not assume any prior data distribution, making it well-suited for non-linear relationships between predictors and the target variable.  
Due to the complex nonlinear relationship between numerical and categorical characteristics in the Airbnb dataset (such as room type, number of amenities, and ratings), this study adopts the k-Nearest Neighbour algorithm (k-NN) as a simple and effective model to identify and learn these potential patterns in the multi-dimensional feature space.  
The algorithm measures the similarity by calculating the distance between samples without assuming a specific pattern in advance, so it can better adapt to the combination of such nonlinear features.

### Data Preprocessing
Since kNN measures similarity by the distance between samples, all input features are converted to numerical types and standardized to the same scale.  
For the categorical variable room_type, one-hot encoding is used to ensure that the model can identify the differences among categories without artificially introducing ordinal relationships.   
In terms of numerical features, in order to improve the stability of the model and reduce the redundant information in distance calculation, we performed a three-step feature screening.     
(1) removing near-zero variance features,  
(2) dropping highly correlated predictors (|ρ| > 0.85), and  
(3) retaining variables with a moderate Spearman correlation (|ρ| > 0.15) with the target price_class.  
Finally, all numerical features are centered and scaled to prevent the characteristics with larger values from having an unbalanced effect in the calculation of Euclidean distance.

```{r}
library(caret)
library(yardstick)
library(reshape2)
set.seed(123)

# 1. split dataset 
df <- df_filtered
df$price_class <- factor(
  df$price_class,
  levels = c("[0,150)", "[150,300)", "[300,+)"),
  ordered = TRUE
)

idx <- createDataPartition(df$price_class, p = 0.8, list = FALSE)
train_df <- df[idx, ]
test_df  <- df[-idx, ]

# 2. categorical feature engineering for kNN

dummies <- dummyVars(~ room_type, data = train_df)

train_room <- predict(dummies, newdata = train_df) %>% as.data.frame()
test_room  <- predict(dummies, newdata = test_df)  %>% as.data.frame()
stopifnot(all(colnames(train_room) == colnames(test_room)))
train_df <- bind_cols(train_df %>% select(-room_type), train_room)
test_df  <- bind_cols(test_df  %>% select(-room_type), test_room)

# 3. Numerical feature selection
num_vars <- train_df %>%
  select(where(is.numeric)) %>%
  select(-price, -price_orig)

# (a) Remove columns with little or no variance (near-zero variance)
low_var <- nearZeroVar(num_vars)
if (length(low_var) > 0) {
  num_vars <- num_vars[, -low_var, drop = FALSE]
}
# (b) Remove columns that are highly colinear

corr_mat <- cor(num_vars, use = "complete.obs")
high_corr <- findCorrelation(corr_mat, cutoff = 0.85)
if (length(high_corr) > 0) {
  num_vars <- num_vars[, -high_corr, drop = FALSE]
}
# (c) Do a final round of screening based on relevance to the target
corr_with_y <- sapply(
  num_vars,
  function(x) cor(
    x,
    as.numeric(train_df$price_class),
    method = "spearman",
    use = "complete.obs"
  )
)
selected_vars <- names(corr_with_y[abs(corr_with_y) > 0.15])

train_filtered <- train_df %>%
  select(all_of(selected_vars), price_class)

test_filtered <- test_df %>%
  select(all_of(selected_vars), price_class)
```

### Model Tuning and Training
The model’s hyperparameter, k (number of neighbours), was tuned using 5-fold cross-validation on the training set. A search range from k = 15 to 31 (odd numbers) was tested, and the best value was selected based on classification accuracy.
After tuning, the final model was retrained on the full training data using the optimal k and subsequently evaluated on the independent 20% test set, which was held out from all model selection procedures.
```{r}
# 4. Parameter tuning
ctrl_inner <- trainControl(method = "cv", number = 5)
grid_k <- expand.grid(k = seq(15, 31, 2)) 
knn_cv <- train(
  price_class ~ .,
  data = train_filtered,
  method = "knn",
  trControl = ctrl_inner,
  preProcess = c("center", "scale"),
  tuneGrid = grid_k,
  metric = "Accuracy"
)

best_k <- knn_cv$bestTune$k

# 5. Refit the final model with the best k
best_knn <- train(
  price_class ~ .,
  data = train_filtered,
  method = "knn",
  trControl = trainControl(method = "none"),
  preProcess = c("center", "scale"),
  tuneGrid = data.frame(k = best_k)
)
cat("Best k from 5-fold CV:", best_k, "\n")
```

## Support Vector Machine (SVM, RBF kernel)
Support Vector Machine (SVM) is a supervised learning algorithm that aims to find the optimal hyperplane separating data points from different classes by maximizing the margin between them.
It is particularly effective for high-dimensional and complex datasets where the relationship between features and labels is not linearly separable.

In this project, we employed the Radial Basis Function (RBF) kernel version of SVM because our Airbnb dataset shows non-linear feature interactions (e.g., the combination of location rating, number of amenities, and reviews affects price category in a non-linear way).
The RBF kernel maps data into a higher-dimensional space, allowing the SVM to capture curved decision boundaries that cannot be represented by a simple linear model.

### Kernel Function
The RBF kernel is defined as:
$$
K(x_i, x_j) = \exp(-\gamma \|x_i - x_j\|^2)
$$

* xi, xj are feature vectors
* γ (gamma) controls how far the influence of a single training sample reaches — a small gamma means a smoother decision boundary, while a large gamma makes the model more sensitive to individual points.

The SVM with RBF kernel essentially transforms the feature space so that data from different classes become linearly separable in the higher-dimensional space.

### Parameter Tuning and Training Process
Before training, the dataset was split into 75% training data and 25% testing data, ensuring class balance across splits and delete some url columns.

```{r, message=FALSE, warning=FALSE, results='hide'}

library(e1071)
library(caret)
set.seed(123)

# --- 1. split dataset ---
df_filtered$price_class <- factor(df_filtered$price_class)
index <- createDataPartition(df_filtered$price_class, p = 0.75, list = FALSE)
svm_train <- df_filtered[index, ]
svm_test  <- df_filtered[-index, ]

# --- 2. remain dependent variables and independent variables ---
vars <- c("accommodates", "bedrooms", "beds", "room_type", "neighbourhood_cleansed", "price_class")
svm_train <- svm_train[, vars]
svm_test  <- svm_test[, vars]
head(svm_train)



```

To find the optimal parameters, we used grid search combined with k-fold cross-validation (k = 5) over a range of values for C and gamma.
Cross-validation helps to reduce overfitting and ensures that the selected parameters generalize well across unseen data. And according to the result the best parameters are C = 1, γ = 1
The final model was then trained on the full training set with the best (C, γ) combination and evaluated on the testing set.
```{r, message=FALSE, warning=FALSE, results='hide'}

# --- 3. make sure no missing value ---
svm_train <- na.omit(svm_train)
svm_test  <- na.omit(svm_test)

# --- 4. room_type and neighbourhood_cleansed  ---
svm_train$room_type <- as.factor(svm_train$room_type)
svm_test$room_type  <- factor(svm_test$room_type, levels = levels(svm_train$room_type))

svm_train$neighbourhood_cleansed <- as.factor(svm_train$neighbourhood_cleansed)
svm_test$neighbourhood_cleansed  <- factor(svm_test$neighbourhood_cleansed, levels = levels(svm_train$neighbourhood_cleansed))


```

```{r, message=FALSE, warning=FALSE, eval=FALSE}
# --- 5. Grid search and across validation ---
tuned <- tune(
  svm,
  price_class ~ .,
  data = svm_train,
  kernel = "radial",
  ranges = list(cost = c(0.1, 1, 10), gamma = c(0.01, 0.1, 1)),
  tunecontrol = tune.control(cross = 3)
)


```


```{r, message=FALSE, warning=FALSE, results='hide'}
# --- 6. output optimal parameters ---
best_model <- svm(
  price_class ~ .,
  data = svm_train,
  kernel = "radial",
  cost = 1,
  gamma = 1
)


# --- 7. predict ---
svm_pred<- predict(best_model, newdata = svm_test)
```




## Random Forest

###Model Overview

Random Forest is an ensemble of decision trees trained on bootstrap samples with random feature subsampling at each split. It reduces variance while preserving non-linear decision boundaries and high-order interactions. It requires minimal distributional assumptions and works well with mixed numeric–categorical predictors after one-hot encoding.

###Data Preprocessing

Target: price_class as an ordered factor with three levels [0,150), [150,300), [300,+).

Numeric: replace sentinel -1 with within-fold median (computed on the training split only).

Categorical: one-hot encode; for high-cardinality factors keep top 10 levels and collapse the rest to Other.

Leakage control: all steps executed inside each CV fold.
```{r}

## --- Random Forest 
library(tidyverse)
library(caret)
library(ranger)
library(pROC)
library(knitr)

set.seed(123)

# --- 1. Load data ---
df_clean <- read.csv("cleaned_data_with_price_class_no_dislocation.csv")

# --- 2. Encode target variable ---
df_clean$price_class <- trimws(df_clean$price_class)
df_clean <- df_clean[df_clean$price_class %in% c("[0,150)", "[150,300)", "[300,+)"), ]

df_clean$price_class <- factor(df_clean$price_class,
  levels = c("[0,150)", "[150,300)", "[300,+)"),
  labels = c("low", "mid", "high")
)

df_clean <- df_clean[!is.na(df_clean$price_class), ]

cat("\n✅ Class distribution:\n")
print(table(df_clean$price_class))

# --- 3. Remove leakage columns ---
leak_cols <- intersect(
  names(df_clean),
  c("price", "price_orig", "price_usd", "price_clean",
    "price_class_num", "price_bucket")
)
if (length(leak_cols) > 0) {
  message("Removing leakage columns: ", paste(leak_cols, collapse = ", "))
  df_clean <- df_clean[, setdiff(names(df_clean), leak_cols), drop = FALSE]
}

# --- 4. Helper functions ---
f1_macro <- function(truth, pred) {
  lv <- levels(truth)
  f1s <- sapply(lv, function(l) {
    p <- ifelse(pred == l, 1, 0)
    t <- ifelse(truth == l, 1, 0)
    tp <- sum(p == 1 & t == 1)
    fp <- sum(p == 1 & t == 0)
    fn <- sum(p == 0 & t == 1)
    prec <- ifelse(tp + fp == 0, 0, tp / (tp + fp))
    rec  <- ifelse(tp + fn == 0, 0, tp / (tp + fn))
    if ((prec + rec) == 0) 0 else 2 * prec * rec / (prec + rec)
  })
  mean(f1s)
}

mcc_binary <- function(t, p) {
  tp <- sum(p == 1 & t == 1)
  tn <- sum(p == 0 & t == 0)
  fp <- sum(p == 1 & t == 0)
  fn <- sum(p == 0 & t == 1)
  denom <- sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
  if (is.na(denom) || denom == 0) return(NA_real_)
  (tp * tn - fp * fn) / denom
}

mcc_macro <- function(truth, pred) {
  lv <- levels(truth)
  mccs <- sapply(lv, function(l) {
    if (!any(truth == l)) return(NA_real_)
    t <- ifelse(truth == l, 1, 0)
    p <- ifelse(pred == l, 1, 0)
    mcc_binary(t, p)
  })
  mean(mccs, na.rm = TRUE)
}

# --- 5. Stratified 5-fold ---
folds <- createFolds(df_clean$price_class, k = 5, list = TRUE)

rf_fold_metrics <- data.frame(Fold = integer(),
                              Accuracy = double(),
                              MacroF1 = double()
                              )
rf_fold_preds <- list()
rf_fold_truth <- list()

#----
mcc_binary <- function(t, p) {
  tp <- sum(p == 1 & t == 1)
  tn <- sum(p == 0 & t == 0)
  fp <- sum(p == 1 & t == 0)
  fn <- sum(p == 0 & t == 1)
  denom <- sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
  if (is.na(denom) || denom == 0) return(0)
  (tp * tn - fp * fn) / denom
}

mcc_macro <- function(truth, pred) {
  lv <- levels(truth)
  mccs <- sapply(lv, function(l) {
    t <- ifelse(truth == l, 1, 0)
    p <- ifelse(pred == l, 1, 0)
    mcc_binary(t, p)
  })
  mean(mccs, na.rm = TRUE)
}

fold_summary <- character()



# --- 6. Cross-validation loop (修正版) ---
for (i in seq_along(folds)) {
  test_idx <- folds[[i]]
  train_idx <- setdiff(seq_len(nrow(df_clean)), test_idx)
  tr <- df_clean[train_idx, , drop = FALSE]
  te <- df_clean[test_idx, , drop = FALSE]

  # skip if class missing
  if (length(unique(tr$price_class)) < length(levels(df_clean$price_class))) {
    cat(sprintf("⚠️ Fold %d skipped (missing class)\n", i))
    next
  }

  # train
  rf_fit <- ranger(
    price_class ~ ., data = tr,
    num.trees = 500,
    importance = "impurity",
    probability = TRUE
  )

  # predict
  pr <- predict(rf_fit, te, type = "response")

  
  prob_mat <- pr$predictions[, levels(tr$price_class), drop = FALSE]

  pred_labels <- colnames(prob_mat)[max.col(prob_mat, ties.method = "first")]

 
  pr_cls <- factor(pred_labels, levels = levels(tr$price_class))
  truth <- te$price_class

  # --- metrics ---
  acc <- mean(pr_cls == truth)
  f1m <- f1_macro(truth, pr_cls)
  mccm <- mcc_macro(truth, pr_cls)

  rf_fold_preds[[i]] <- pr_cls
  rf_fold_truth[[i]] <- truth
  rf_fold_metrics <- rbind(rf_fold_metrics,
                           data.frame(Fold = i,
                                      Accuracy = acc,
                                      MacroF1 = f1m
                                      ))
}

#cat("\n5-fold cross-validation completed.\n")
kable(rf_fold_metrics, caption = "5-Fold Cross Validation Results for Random Forest",
      digits = 4)

# --- 7. Results summary ---
summ <- rf_fold_metrics |>
  summarise(
    Accuracy = sprintf("%.4f ± %.4f", mean(Accuracy), sd(Accuracy)),
    MacroF1  = sprintf("%.4f ± %.4f", mean(MacroF1), sd(MacroF1)),
    
  )

cat("\n📊 Performance Metrics (Mean ± SD):\n")
print(summ)

# --- 8. Aggregated confusion matrix ---
label_map <- c("low" = "[0,150)", "mid" = "[150,300)", "high" = "[300,+)")
all_pred  <- factor(unlist(rf_fold_preds), levels = names(label_map), labels = label_map)
all_truth <- factor(unlist(rf_fold_truth), levels = names(label_map), labels = label_map)

cm <- confusionMatrix(all_pred, all_truth)
cat("\nAggregated Confusion Matrix:\n")
print(cm$table)
cat(sprintf("\nOverall Accuracy: %.4f\n", mean(all_pred == all_truth)))

# --- 9. Feature importance ---
rf_full <- ranger(
  price_class ~ ., data = df_clean,
  num.trees = 500,
  importance = "impurity",
  probability = TRUE
)

imp <- sort(rf_full$variable.importance, decreasing = TRUE)
top_imp <- tibble(Feature = names(imp)[1:min(20, length(imp))],
                  Importance = as.numeric(imp[1:min(20, length(imp))]))
cat("\n🏆 Top 20 Feature Importances:\n")
print(top_imp)

rf_cls <- factor(unlist(rf_fold_preds), levels = levels(df_clean$price_class))
rf_test_df <- data.frame(price_class = unlist(rf_fold_truth))

```











# Performance Evaluation
## Softmax Logistic Regression
### Cross-Validation Results Summary
The cross-validation results are aggregated to compute mean performance and variability measures. Standard deviations quantify the stability of model performance across different data splits, while 95% confidence intervals provide a range for the expected performance on unseen data.

```{r, message=FALSE, warning=FALSE}
# Calculate summary statistics
summary_table <- data.frame(
  Metric = c("Accuracy", "Macro-F1", "MCC", "Macro-AUC", "Log Loss"),
  Mean = round(c(mean(cv_results$accuracy), mean(cv_results$macro_f1),
                 mean(cv_results$mcc), mean(cv_results$macro_auc),
                 mean(cv_results$logloss)), 4),
  SD = round(c(sd(cv_results$accuracy), sd(cv_results$macro_f1),
               sd(cv_results$mcc), sd(cv_results$macro_auc),
               sd(cv_results$logloss)), 4)
)

summary_table$CI_Lower <- round(summary_table$Mean - 1.96 * summary_table$SD / sqrt(k_folds), 4)
summary_table$CI_Upper <- round(summary_table$Mean + 1.96 * summary_table$SD / sqrt(k_folds), 4)

cat("\nPerformance Metrics (Mean ± SD [95% CI]):\n")
for (i in 1:nrow(summary_table)) {
  cat(sprintf("%s: %.4f ± %.4f [%.4f, %.4f]\n",
              summary_table$Metric[i], summary_table$Mean[i],
              summary_table$SD[i], summary_table$CI_Lower[i],
              summary_table$CI_Upper[i]))
}

cat("\nDetailed Fold-wise Results:\n")
print(knitr::kable(cv_results, digits = 4))
```

### Aggregated Confusion Matrix
An aggregated confusion matrix is constructed by combining predictions from all cross-validation folds. This provides a comprehensive view of the model's classification patterns and reveals which price classes are most frequently confused.

```{r, message=FALSE, warning=FALSE}
# Combine all predictions from CV folds
all_pred <- do.call(c, all_predictions)
all_true <- do.call(c, all_true_labels)

# Ensure both are factors with same levels
all_pred <- factor(all_pred, levels = c("[0,150)", "[150,300)", "[300,+)"))
all_true <- factor(all_true, levels = c("[0,150)", "[150,300)", "[300,+)"))

# Generate aggregated confusion matrix
cm_aggregated <- confusionMatrix(all_pred, all_true)

cat("\nAggregated Confusion Matrix (All Folds):\n")
print(cm_aggregated$table)

cat("\nOverall Accuracy:", round(cm_aggregated$overall['Accuracy'], 4), "\n")

# Display class-wise performance metrics
cat("\nClass-wise Performance:\n")
class_metrics <- cm_aggregated$byClass[, c("Sensitivity", "Specificity", "Precision", "F1")]
print(round(class_metrics, 4))
```

## Confusion Matrix Visualization
The confusion matrix is visualized as a heatmap to facilitate interpretation of classification patterns. Cell colors indicate prediction frequencies, with darker shades representing higher counts.

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
# Prepare data for visualization
cm_df <- as.data.frame(cm_aggregated$table)
names(cm_df) <- c("Prediction", "Reference", "Freq")

# Create heatmap
p1 <- ggplot(cm_df, aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile(color = "white", size = 1.5) +
  geom_text(aes(label = Freq), size = 6, fontface = "bold") +
  scale_fill_gradient(low = "#f0f0f0", high = "#2c7bb6") +
  labs(title = "Confusion Matrix - Softmax Logistic Regression",
       subtitle = "Aggregated from 5-Fold Cross-Validation") +
  theme_minimal(base_size = 14) +
  theme(panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

print(p1)
```

### Cross-Validation Performance Distribution
The distribution of performance metrics across folds is visualized using boxplots. Red diamonds indicate mean values, while boxes show the interquartile range. This visualization reveals the stability and variability of model performance.

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=6}
# Prepare data for visualization
cv_long <- cv_results %>%
  select(fold, accuracy, macro_f1, mcc, macro_auc) %>%
  pivot_longer(cols = -fold, names_to = "metric", values_to = "value")

# Create boxplot
p2 <- ggplot(cv_long, aes(x = metric, y = value, fill = metric)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5, size = 2) +
  stat_summary(fun = mean, geom = "point", shape = 23, 
               size = 3, fill = "red", color = "darkred") +
  labs(title = "Cross-Validation Performance Distribution",
       subtitle = "5-Fold Stratified CV (Red diamonds = Mean)",
       x = "Metric", y = "Value") +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 30, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

print(p2)
```

### Feature Importance Analysis
Feature importance is assessed by computing the average absolute coefficient magnitude across all price classes. Higher values indicate features that have stronger influence on price classification. The top 20 most important features are visualized to identify key predictors.

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=8}
# Train a final model on all data for feature importance extraction
set.seed(42)
train_idx <- createDataPartition(df$price_class, p = 0.8, list = FALSE)
train_raw <- df[train_idx, ]
test_raw <- df[-train_idx, ]

# Preprocess
preprocessed <- preprocess_for_modeling(train_raw, test_raw, 
                                        valid_numeric, key_categorical)

# Train model
train_data <- cbind(target = preprocessed$y_train, preprocessed$X_train)
model_final <- multinom(target ~ ., data = train_data, 
                        maxit = 200, trace = FALSE, MaxNWts = 10000)

# Extract coefficients
coef_matrix <- summary(model_final)$coefficients
feature_names <- rownames(coef_matrix)
feature_names <- feature_names[feature_names != "(Intercept)"]

# Calculate average absolute coefficient across classes
avg_abs_coef <- apply(abs(coef_matrix[feature_names, ]), 1, mean)
importance_df <- data.frame(
  feature = feature_names,
  importance = avg_abs_coef
) %>% arrange(desc(importance)) %>% head(20)

# Create bar plot
p3 <- ggplot(importance_df, aes(x = reorder(feature, importance), y = importance)) +
  geom_col(fill = "#d95f02", alpha = 0.8) +
  coord_flip() +
  labs(title = "Top 20 Feature Importance",
       subtitle = "Average |Coefficient| across Price Classes",
       x = "Feature", y = "Importance") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

print(p3)

cat("\nTop 10 Most Important Features:\n")
print(knitr::kable(head(importance_df, 10), digits = 4, row.names = FALSE))
```

### Performance Summary and Interpretation
The softmax logistic regression model demonstrates stable performance across cross-validation folds, providing a reliable baseline for comparison with more complex models. The confusion matrix reveals the model's classification patterns, showing where predictions are most accurate and where common misclassifications occur. Feature importance analysis identifies the key drivers of price classification, with property characteristics and location features playing prominent roles.

```{r, message=FALSE, warning=FALSE}
cat("\n", rep("=", 70), "\n")
cat("SOFTMAX LOGISTIC REGRESSION - FINAL SUMMARY\n")
cat(rep("=", 70), "\n\n")

cat("Model Specification:\n")
cat("  Algorithm:      Multinomial Logistic Regression (Softmax)\n")
cat("  Features:      ", ncol(preprocessed$X_train), "\n")
cat("  CV Strategy:    5-Fold Stratified Cross-Validation\n")
cat("  Total Samples: ", nrow(df), "\n\n")

cat("CROSS-VALIDATION PERFORMANCE:\n")
cat(sprintf("  Accuracy:    %.4f ± %.4f (Mean ± SD)\n", 
            mean(cv_results$accuracy), sd(cv_results$accuracy)))
cat(sprintf("  Macro-F1:    %.4f ± %.4f\n", 
            mean(cv_results$macro_f1), sd(cv_results$macro_f1)))
cat(sprintf("  MCC:         %.4f ± %.4f\n", 
            mean(cv_results$mcc), sd(cv_results$mcc)))
cat(sprintf("  Macro-AUC:   %.4f ± %.4f\n", 
            mean(cv_results$macro_auc), sd(cv_results$macro_auc)))
cat(sprintf("  Log Loss:    %.4f ± %.4f\n\n", 
            mean(cv_results$logloss), sd(cv_results$logloss)))

cat("95% CONFIDENCE INTERVALS:\n")
for (i in 1:nrow(summary_table)) {
  cat(sprintf("  %s: [%.4f, %.4f]\n",
              summary_table$Metric[i], 
              summary_table$CI_Lower[i],
              summary_table$CI_Upper[i]))
}

cat("\nKEY FINDINGS:\n")
cat("✓ Model exhibits", 
    ifelse(sd(cv_results$accuracy) < 0.05, "highly stable", "stable"), 
    "performance across folds\n")
cat("✓ Macro-averaged metrics ensure balanced evaluation across all price classes\n")
cat("✓ Feature importance analysis reveals interpretable predictors\n")
cat("✓ Ready for comparison with alternative classification algorithms\n")

cat("\n", rep("=", 70), "\n")
```

## Linear Discriminant Analysis (LDA)
```{r, message=FALSE, warning=FALSE, results='hide'}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)

file <- "cleaned_data_with_price_class_no_dislocation.csv"
target_cols <- c("price_class", "room_type", "neighbourhood_cleansed")

df_selected <- read_csv(
  file,
  col_select = all_of(target_cols)
)

df_lda <- df_selected %>%
    mutate(
        price_class = factor(price_class),
        room_type = factor(room_type),
        neighbourhood_cleansed = factor(neighbourhood_cleansed)
    )


df_after <- df_lda %>%
    drop_na(price_class, room_type, neighbourhood_cleansed)

library(MASS)
library(ggplot2)

lda_fit <- lda(price_class ~ room_type + neighbourhood_cleansed, data = df_after)
lda_pred <- predict(lda_fit, newdata = df_after)

tab <- table(Predicted = lda_pred$class, True = df_after$price_class)
acc <- sum(diag(tab)) / sum(tab)

ld_df <- data.frame(lda_pred$x, price_class = df_after$price_class)
ggplot(ld_df, aes(x = LD1, y = LD2, color = price_class)) +
  geom_point(size = 2, alpha = 0.7) +
  theme_minimal() +
  labs(title = "LDA projection (LD1 vs LD2)",
       x = "LD1", y = "LD2")
tab
```
* Overall accuracy on the evaluated set is **", round(acc,3), "**.
* Macro-F1 (unweighted average of per-class F1) is **", round(macroF1,3), "**.
* Multiclass MCC is **", ifelse(is.na(mcc_val), "NA", round(mcc_val,3)), "**.
  These metrics indicate **moderate discriminative performance**: the model classifies correctly substantially more often than chance, but there remains a meaningful error rate (~40% misclassification).

The confusion matrix shows that the middle class `[150,300)` is the largest class and is often predicted correctly, but also causes many misclassifications in both directions (predicted as either lower or higher price class). Conversely, the lowest class `[0,150)` is sometimes over-predicted and the highest class `[300,+)` is partly confused with the middle class. This pattern is typical when the target classes are adjacent price bands whose boundaries are inherently fuzzy.

LDA provides interpretable linear combinations (coefficients) that indicate how specific neighbourhoods and room type levels move observations along discriminant axes. The printed discriminant coefficients give direction and relative magnitude: positive/negative loadings show which features push listings toward higher vs lower discriminant scores. This supports **qualitative** conclusions about the influence of neighbourhood and room type. However, because many important features are missing from the model and because LDA assumptions are not perfectly met, **quantitative causal claims or precise effect sizes** based solely on this LDA should be made cautiously.

LDA with `room_type` and `neighbourhood_cleansed` captures meaningful signal and yields interpretable discriminant directions, but its predictive performance is moderate because of limited predictors and assumption violations. Let's see how several other models perform on this classification problem.


## K Nearest Neighbour (kNN)
After tuning the neighborhood size (k = 27), the final kNN model was evaluated on a 20% independent test set. 
The results show that the overall accuracy of the model is 0.7079, that is, about 71% of listings are correctly classified into the corresponding price range. 
The macro-weighted F1 score (0.7087) and MCC (0.5507) also show that the model's performance in three price categories was relatively stable, maintaining a good balance between precision and recall rate.
```{r}
# 1. Independent test set evaluation

knn_test_pred <- predict(best_knn, newdata = test_filtered)


knn_acc <- mean(knn_test_pred == test_filtered$price_class)


cm <- confusionMatrix(
  data      = knn_test_pred,
  reference = test_filtered$price_class
)

results_df <- data.frame(
  truth    = test_filtered$price_class,
  estimate = knn_test_pred
)

f1_macro <- results_df %>%
  f_meas(truth = truth, estimate = estimate, estimator = "macro_weighted")


mcc_val <- results_df %>%
  mcc(truth = truth, estimate = estimate)
cat("Test Accuracy:", knn_acc, "\nMacro F1 (weighted):", f1_macro$.estimate, "\nMCC:", mcc_val$.estimate, "\n")



# 2. Visualisation: Confusion Matrix Heat Map
cm_table <- as.data.frame(cm$table)
colnames(cm_table) <- c("Prediction", "Reference", "Freq")

ggplot(cm_table, aes(x = Reference, y = Prediction, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), size = 4, color = "black") +
  scale_fill_gradient(low = "#fbe9e7", high = "#d84315") +
  labs(
    title = paste0("Confusion Matrix — kNN (k = ", best_k, ")"),
    x = "True Label (Reference)",
    y = "Predicted Label"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

As shown in the confusion matrix above, most listings in the middle price range [150,300) were correctly classified, and there is still a certain degree of overlap between adjacent categories.  
Some low-priced listings [0,150) were predicted as mid-priced listings, and a small number of high-priced listings [300,+) were also predicted as mid-priced housing.  
This situation is understandable, because listings close to the price boundary usually have similar characteristics in terms of room type and number of amenities.  
Overall, the kNN model shows stable performance, which can effectively identify listings with similar characteristics, and keeps most misclassifications within adjacent price ranges.

## Support Vector Machine (SVM, RBF kernel)
The SVM classifier with an RBF kernel achieved an overall accuracy of 70.1%, indicating a reasonably strong predictive performance across the three price classes. The macro-averaged precision (0.72) and recall (0.69) suggest that the model performs consistently across all categories, without being overly biased toward the majority class. The F1-score (0.70) further confirms a balanced trade-off between precision and recall. As shown in Figure X, most predictions align with the diagonal in the confusion matrix, especially for the middle price range ([150,300)), implying that the model captures the main pricing patterns effectively.However, some misclassification occurs between adjacent ranges (e.g., low vs. mid, mid vs. high).


```{r message=FALSE, warning=FALSE, results='hide'}
conflicts_prefer(yardstick::precision)


# yardstick 
res <- tibble(
  truth = svm_test$price_class,
  .pred_class = factor(svm_pred, levels = levels(svm_test$price_class))
)

svm_acc   <- accuracy(res, truth, .pred_class)
svm_prec  <- precision(res, truth, .pred_class, estimator = "macro")
svm_rec   <- yardstick::recall(res, truth, .pred_class, estimator = "macro")
svm_f1    <- f_meas(res, truth, .pred_class, estimator = "macro")

metrics_table <- tibble(
  Metric = c("Accuracy", "Precision (Macro)", "Recall (Macro)", "F1-Score (Macro)"),
  Value  = c(svm_acc$.estimate, svm_prec$.estimate, svm_rec$.estimate, svm_f1$.estimate)
)

metrics_table




```

```{r}
library(caret)
library(ggplot2)

cm <- confusionMatrix(svm_pred, svm_test$price_class)

cm$table

cm_df <- as.data.frame(cm$table)

ggplot(cm_df, aes(x = Reference, y = Prediction, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), color = "black", size = 4) +
  scale_fill_gradient(low = "#D6EAF8", high = "#1B4F72") +
  labs(
    title = "Confusion Matrix of SVM (RBF Kernel)",
    x = "True Class",
    y = "Predicted Class",
    fill = "Count"
  ) +
  theme_minimal(base_size = 13)



```


## Random Forest
```{r}
# --- 10. Visualizations ---

library(ggplot2)
library(reshape2)

## (a) Confusion Matrix Heatmap ----
cm_df <- as.data.frame(cm$table)
colnames(cm_df) <- c("Prediction", "Reference", "Freq")

ggplot(cm_df, aes(x = Reference, y = Prediction, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), color = "black", size = 4) +
  scale_fill_gradient(low = "#FEE0D2", high = "#DE2D26") +
  labs(
    title = "Confusion Matrix — Random Forest",
    x = "True Label (Reference)",
    y = "Predicted Label",
    fill = "Freq"
  ) +
  theme_minimal(base_size = 14)

## (b) Feature Importance Bar Plot ----
top_imp_plot <- top_imp[1:15, ]  # 取前15个主要特征
ggplot(top_imp_plot, aes(x = reorder(Feature, Importance), y = Importance)) +
  geom_col(fill = "#3182BD") +
  coord_flip() +
  labs(
    title = "Top Feature Importances — Random Forest",
    x = "Feature",
    y = "Importance (Impurity-based)"
  ) +
  theme_minimal(base_size = 14)


```

```{r rf_performance_table, message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
library(kableExtra)
library(yardstick)

suppressMessages({
  conflicted::conflicts_prefer(yardstick::precision)
  conflicted::conflicts_prefer(yardstick::recall)
  conflicted::conflicts_prefer(yardstick::f_meas)
})

# Calculate key metrics (macro-averaged)
rf_results <- tibble(
  truth = rf_test_df$price_class,
  estimate = rf_cls 
)

rf_acc  <- confusionMatrix(rf_cls , rf_test_df$price_class)$overall["Accuracy"]
rf_prec <- precision(rf_results, truth, estimate, estimator = "macro")
rf_rec  <- recall(rf_results, truth, estimate, estimator = "macro")
rf_f1   <- f_meas(rf_results, truth, estimate, estimator = "macro")
rf_mcc  <- mcc(rf_results, truth, estimate)

# Combine into a summary table
rf_table <- data.frame(
  Metric = c("Accuracy", "Precision (Macro)", "Recall (Macro)", "F1-Score (Macro)", "MCC"),
  Value  = c(rf_acc, rf_prec$.estimate, rf_rec$.estimate, rf_f1$.estimate, rf_mcc$.estimate)
)

# Display as formatted table
kable(rf_table, digits = 4, caption = "Random Forest Model Performance Summary") %>%
  kable_styling(full_width = FALSE, position = "center",
                bootstrap_options = c("striped", "hover", "condensed"))

```



Across folds, accuracy and macro-F1 are closely aligned, suggesting balanced performance across classes rather than dominance by the majority [150,300) band. Macro-MCC confirms consistent correlation between predictions and ground truth under class imbalance. The aggregated confusion matrix shows most errors occur at the boundaries between [0,150) and [150,300), and between [150,300) and [300,+), consistent with fuzzy price thresholds. Feature-importance results corroborate earlier EDA: capacity and neighbourhood signals jointly structure the price tiers.




## Comparison of Models

```{r fill_missing_metrics, message=FALSE, warning=FALSE}

## ====== Fill missing values in comparison_df ======

# --- Compute missing Softmax precision / recall ---
softmax_prec <- mean(sapply(levels(all_true), function(l) {
  tp <- sum(all_pred == l & all_true == l)
  fp <- sum(all_pred == l & all_true != l)
  if (tp + fp == 0) 0 else tp / (tp + fp)
}))
softmax_rec <- mean(sapply(levels(all_true), function(l) {
  tp <- sum(all_pred == l & all_true == l)
  fn <- sum(all_pred != l & all_true == l)
  if (tp + fn == 0) 0 else tp / (tp + fn)
}))
softmax_acc  <- mean(cv_results$accuracy, na.rm = TRUE)
softmax_prec <- mean(softmax_prec, na.rm = TRUE)
softmax_rec  <- mean(softmax_rec, na.rm = TRUE)
softmax_f1   <- mean(cv_results$macro_f1, na.rm = TRUE)

svm_prec <- mean(svm_prec, na.rm = TRUE)
svm_rec  <- mean(svm_rec,  na.rm = TRUE)
svm_f1   <- mean(svm_f1,   na.rm = TRUE)
# --- Compute missing LDA precision / recall / F1 / MCC ---
lda_prec <- mean(sapply(levels(df_after$price_class), function(l) {
  tp <- sum(lda_pred$class == l & df_after$price_class == l)
  fp <- sum(lda_pred$class == l & df_after$price_class != l)
  if (tp + fp == 0) 0 else tp / (tp + fp)
}))
lda_rec <- mean(sapply(levels(df_after$price_class), function(l) {
  tp <- sum(lda_pred$class == l & df_after$price_class == l)
  fn <- sum(lda_pred$class != l & df_after$price_class == l)
  if (tp + fn == 0) 0 else tp / (tp + fn)
}))
lda_f1 <- mean(sapply(levels(df_after$price_class), function(l) {
  tp <- sum(lda_pred$class == l & df_after$price_class == l)
  fp <- sum(lda_pred$class == l & df_after$price_class != l)
  fn <- sum(lda_pred$class != l & df_after$price_class == l)
  p <- ifelse(tp + fp == 0, 0, tp / (tp + fp))
  r <- ifelse(tp + fn == 0, 0, tp / (tp + fn))
  if ((p + r) == 0) 0 else 2 * p * r / (p + r)
}))
lda_mcc <- mcc_macro(df_after$price_class, lda_pred$class)

# --- Compute missing KNN precision / recall ---
knn_prec <- mean(sapply(levels(test_filtered$price_class), function(l) {
  tp <- sum(knn_test_pred == l & test_filtered$price_class == l)
  fp <- sum(knn_test_pred == l & test_filtered$price_class != l)
  if (tp + fp == 0) 0 else tp / (tp + fp)
}))
knn_rec <- mean(sapply(levels(test_filtered$price_class), function(l) {
  tp <- sum(knn_test_pred == l & test_filtered$price_class == l)
  fn <- sum(knn_test_pred != l & test_filtered$price_class == l)
  if (tp + fn == 0) 0 else tp / (tp + fn)
}))


knn_f1 <- mean(sapply(levels(test_filtered$price_class), function(l) {
  tp <- sum(knn_test_pred == l & test_filtered$price_class == l)
  fp <- sum(knn_test_pred == l & test_filtered$price_class != l)
  fn <- sum(knn_test_pred != l & test_filtered$price_class == l)
  
  prec <- ifelse(tp + fp == 0, 0, tp / (tp + fp))
  rec  <- ifelse(tp + fn == 0, 0, tp / (tp + fn))
  
  if ((prec + rec) == 0) 0 else 2 * prec * rec / (prec + rec)
}))


```
```{r}
lengths(list(Model, Accuracy, Precision, Recall))

```


```{r}
lda_acc  <- as.numeric(lda_acc)

rf_acc   <- as.numeric(rf_acc)

rf_prec <- as.numeric(precision(rf_results, truth, estimate, estimator = "macro")$.estimate[1])
rf_rec  <- as.numeric(recall(rf_results, truth, estimate, estimator = "macro")$.estimate[1])
rf_f1   <- as.numeric(f_meas(rf_results, truth, estimate, estimator = "macro")$.estimate[1])

svm_prec <- as.numeric(svm_prec)
svm_rec  <- as.numeric(svm_rec)
svm_f1   <- as.numeric(svm_f1)
svm_acc <- as.numeric(confusionMatrix(svm_pred, svm_test$price_class)$overall["Accuracy"][1])

# ==== Build comparison_df safely ====
comparison_df <- base::data.frame(
  Model = c("Softmax Regression", "LDA", "kNN", "SVM (RBF)", "Random Forest"),
  Accuracy = c(
    mean(cv_results$accuracy),
    lda_acc,
    knn_acc,
    svm_acc,   
    rf_acc
  ),
  Precision = c(
    softmax_prec,
    lda_prec,
    knn_prec,
    svm_prec,   
    rf_prec
  ),
  Recall = c(
    softmax_rec,
    lda_rec,
    knn_rec,
    svm_rec,    
    rf_rec
  ),
  F1_Score = c(
    mean(cv_results$macro_f1),
    lda_f1,
    knn_f1,
    svm_f1,     
    rf_f1
  ),
  stringsAsFactors = FALSE
)

# === Format numeric columns ===
comparison_df[, 2:5] <- round(comparison_df[, 2:5], 4)

# === Pretty print as table ===
kable(
  comparison_df,
  caption = "Performance Comparison Across Models",
  align = "c"
) %>%
  kable_styling(
    full_width = FALSE,
    position = "center",
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  column_spec(1, bold = TRUE) %>%
  add_header_above(c(" " = 1, "Performance Metrics" = 4))

```







Across all five classification algorithms — Softmax Regression, Linear Discriminant Analysis (LDA), k-Nearest Neighbors (kNN), Support Vector Machine (SVM with RBF kernel), and Random Forest — distinct trade-offs emerge between predictive accuracy, generalization ability, and interpretability.

Softmax Regression achieves the highest overall performance, with Accuracy = 0.918 and F1 = 0.919. Its balanced Precision (0.920) and Recall (0.919) indicate stable classification across all price tiers. This suggests that the linear decision boundaries it learns are sufficient to capture the primary structure in the data, making it both effective and interpretable.

LDA performs moderately, with Accuracy = 0.598 and F1 = 0.595, consistent with its simplifying Gaussian and equal-covariance assumptions. Its Precision (0.625) and Recall (0.582) remain relatively balanced, suggesting that it still identifies meaningful linear separations despite reduced flexibility compared to nonlinear models.

kNN yields Accuracy = 0.708 and F1 = 0.709, striking a reasonable balance between local adaptivity and noise sensitivity. Its Precision (0.728) and Recall (0.700) indicate that it captures neighborhood-based structure effectively, though performance can vary with data density.

SVM (RBF) performs comparably with Accuracy = 0.701 and F1 = 0.704. The Precision (0.720) and Recall (0.694) values imply solid boundary discrimination and moderate generalization. This model benefits from nonlinear flexibility but may require tuning of kernel parameters for further gains.

Random Forest shows strong improvement compared to earlier runs, achieving Accuracy = 0.802 and F1 = 0.805. Its Precision (0.820) and Recall (0.797) reflect robust ensemble learning and better handling of nonlinearity and mixed data types. This suggests it generalizes well while remaining resilient to outliers.

Overall, Softmax Regression still delivers the most consistent and interpretable results, while Random Forest demonstrates powerful generalization with nonlinear interactions. kNN and SVM (RBF) achieve competitive middle-ground performance, and LDA offers a simple yet informative linear baseline.


# Conclusion


# Reference

Brier, G. W. (1950). Verification of forecasts expressed in terms of probability. Monthly Weather Review, 78(1), 1–3.*

Cawley, G. C., & Talbot, N. L. C. (2010). On over-fitting in model selection and subsequent selection bias in performance evaluation. Journal of Machine Learning Research, 11, 2079–2107.

Chicco, D., & Jurman, G. (2020). The advantages of the Matthews correlation coefficient (MCC) over F1 score and accuracy in binary classification evaluation. BMC Genomics, 21, 6. https://doi.org/10.1186/s12864-019-6413-7

Demšar, J. (2006). Statistical comparisons of classifiers over multiple data sets. Journal of Machine Learning Research, 7, 1–30.

Dash, C. S. K., Behera, A. K., Dehuri, S., & Ghosh, A. (2023). An outliers detection and elimination framework in classification task of data mining. Decision Analytics Journal, 6, 100164.

Guo, C., Pleiss, G., Sun, Y., & Weinberger, K. Q. (2017). On calibration of modern neural networks. In Proceedings of the 34th International Conference on Machine Learning (ICML) (pp. 1321–1330).

Hand, D. J., & Till, R. J. (2001). A simple generalisation of the area under the ROC curve for multiple class classification problems. Machine Learning, 45(2), 171–186. https://doi.org/10.1023/A:1010920819831

Lin, H., Zhang, R., & Tong, T. (2025). When Tukey meets Chauvenet: A new boxplot criterion for outlier detection. Journal of Computational and Graphical Statistics.

Mekelburg, E., & Strauss, J. (2024). Pooling and winsorizing machine learning forecasts to predict stock returns with high-dimensional data. Journal of Empirical Finance, 79, 101538.

Nadeau, C., & Bengio, Y. (2003). Inference for the generalization error. Machine Learning, 52(3), 239–281. https://doi.org/10.1023/A:1024068626366

Powers, D. M. W. (2011). Evaluation: From precision, recall and F-measure to ROC, informedness, markedness & correlation. Journal of Machine Learning Technologies, 2(1), 37–63.
